\textls[-0]{\thetitle}

### 2023-03-04 casestudies.tex
make all comments into pdf comment from reviewer.
resolve most comments

Our second case study is a mechanization of abstract interpreters
for simple imperative languages.
In addition to the metatheories, this case study produces abstract
interpreters that are directly ready for program extraction.

@@

Our implementation of family polymorphism retains the computation
ability of the host proof assistant;
the two abstract interpreters in \lsti{ImpTI} and \lsti{ImpCP} are
ready for program extraction to OCaml.
We put them to test over queries.

@@

In addition to the two case studies above, we also use extensible
inductive types for modeling extensible context-free grammars and derive
decision procedures for them.
%\EDJ{I will suggest change to ``and derive decision procedures for them (based on naive parsing by enumeration)''.}

>>>>

Our second case study is a mechanization of abstract interpreters
for simple imperative languages.
In addition to a soundness proof, this case study produces abstract
interpreters that are directly ready for program extraction.

@@

Our implementation of family polymorphism is compatible with Coq's
program extraction feature.
We extract the two verified abstract interpreters to OCaml
and put them to test over simple queries. %\rD{What were the results of testing the extracted abstract interpreters? I don't expect much in the way of performance, but did they return the correct results? What sorts of benchmarks did you use?}

@@

In addition to the two case studies above, we also use extensible
inductive types for modeling extensible context-free grammars and derive
decision procedures for language membership.
%\EDJ{I will suggest change to ``and derive decision procedures for them (based on naive parsing by enumeration)''.}

### 2023-03-04 challenge.tex
make all comments into pdf comment from reviewer.
resolve most comments

\noindentparagraph{%
\headinglabel{C1}{chg:extensible-v-exhaustive}.\,
Extensible inductive types vs.\ exhaustive inductive reasoning.% Reviewer D
}

@@

This flexibility, however, prevents \emph{definitional equality}
that one takes for granted when programming in a proof assistant.
In \cref{fig:stlc-nonmechanized}, a proof assistant supporting family
polymorphism cannot unfold references to the substitution function into its definition (i.e., a pattern match against four cases), %Review D : l135
as a derived family may modify the definition of the function by adding
new cases.
Without the ability to unfold the substitution function, how can the
programmer even prove the substitution lemma?
%
The problem is compounded by the occasional need in derived families to override fields,
which is potentially at odds with being able to use equalities over the fields.

>>>>

\noindentparagraph{%
\headinglabel{C1}{chg:extensible-v-exhaustive}.\,
Extensible inductive types vs.\ exhaustive inductive reasoning.
}

@@

This flexibility, however, prevents \emph{definitional equality}
that one takes for granted when programming in a proof assistant.
In \cref{fig:stlc-nonmechanized}, a proof assistant supporting family
polymorphism cannot unfold references to the substitution function into a
pattern match against four cases,
as a derived family may modify the definition of the function by adding
new cases.
Without the ability to unfold the substitution function, how can the
programmer even prove the substitution lemma?
%
The problem is compounded by the occasional need in derived families to override fields,
which is potentially at odds with being able to use equalities over the fields.

### 2023-03-04 coqimpl.tex
make all comments into pdf comment from reviewer.
resolve most comments

Consider the translation of \lsti{subst} in family \lsti{STLC}.
First, a module named \lsti{STLC\_subst\_Cases} is generated interactively:
every time the programmer completes a \lsti{Case}, the case handler (e.g., \lsti{Def subst\_tm_unit})
is type-checked and added to the module. % Reviewer D: l487

>>>>

Consider the translation of \lsti{subst} in family \lsti{STLC}.
First, a module named \lsti{STLC\_subst\_Cases} is generated on the fly.
%every time the programmer completes a \lsti{Case}, the case handler (e.g., \lsti{Def subst\_tm_unit})
%is type-checked and added to the module. %\rD{This is also confusing-- why is the programmer involved in the translation? Is it that modules are being generated in the background, while commands are being processed?}\EDJreply{changed to mention "type-checked" explicitly. please review}
Importantly, programming remains interactive, as the programmer need not wait until the
entire \lsti{FRecursion} definition is completed to have a \lsti{Case} command
type-checked.

### 2023-03-04 lang-design.tex
resolve most comments

\cref{fig:stlc-mechanized} shows how STLC and its extension with
fixpoints can be mechanized using \Lang, in a style envisioned in
\cref{fig:stlc-nonmechanized}.
The base family "STLC" hosts the STLC metatheories, from abstract syntax
to the type-safety theorem.
Family "STLCFix", derived from "STLC", makes adjustments as needed by a fixpoints extension:
it adds new constructors to the inductive types (\lsti{FInductive}) and
adds new cases to the recursive functions (\lsti{FRecursion}) and
induction proofs (\lsti{FInduction}).
Existing constructors and cases, as well as those definitions and
theorems that need no adjustments ("ty", "env", "empty", "steps", and
"typesafe"), are automatically inherited and reused.
In particular, executing the last command, \lsti{Check STLCFix.typesafe},
displays the type-safety theorem of the fixpoints extension.

@@

In family "STLCFix", inductive type "tm" \emph{further binds} the "tm" type
in family "STLC". It has five constructors: four inherited from
"STLC" and a fifth called "tm_fix".

@@

With inductive types made extensible, exhaustivity of induction becomes
in jeopardy, however.
In particular, recursor "tm_rect" should no longer be allowed, because
its type mentions "tm", whose meaning is late bound, yet
"tm_rect" purports to claim \lsti{\forall t, P t} given only four case handlers.

>>>>

\cref{fig:stlc-mechanized} shows how STLC and its extension with
fixpoints can be mechanized using \Lang, in a style envisioned in
\cref{fig:stlc-nonmechanized}.
%directly adopted from \citet{sf-pl}. \rB{mention non-capture avoiding}\EDJreply{SF is using non-capture avoiding. Is mentioning SF here enough?}\YZreply{It's mentioned in Case Studies. I don't think we need to mention it here.}
The base family "STLC" hosts the STLC metatheories, from abstract syntax
to the type-safety theorem.
Family "STLCFix", derived from "STLC", makes adjustments as needed by a fixpoints extension:
it adds new constructors to the inductive types (\lsti{FInductive}) and
adds new cases to the recursive functions (\lsti{FRecursion}) and
induction proofs (\lsti{FInduction}).
Existing constructors and cases, as well as those definitions and
theorems that need no adjustments ("ty", "env", "empty", "steps", and
"typesafe"), are automatically inherited and reused.
In particular, executing the last command, \lsti{Check STLCFix.typesafe},
displays the type-safety theorem of the fixpoints extension.

@@

In family "STLCFix", %\rD{l175}\EDJreply{Then do we add definite article in front of family "STLCFix"?}
the \lsti{FInductive} "tm" \emph{further binds} the "tm" type
in family "STLC". It has five constructors: four inherited from
"STLC" and a fifth called "tm_fix".

@@

With inductive types made extensible, exhaustivity of induction is now
in question, however.
In particular, the recursor "tm_rect" should no longer be allowed, because
its type mentions "tm", whose meaning is late bound, yet
"tm_rect" purports to claim \lsti{\forall t, P t} given only four case handlers.
